Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-15 paper/paper.tex
another pass on 4.3.4

\subsubsection{Frequency of content modifiers.} As previously
mentioned, \percRulesWithContent\ of Suricata rules involve content
options. Furthermore, we observed---by analyzing the dataset of
existing rules (see Section~\ref{sec:dataset-benign})---that often
these rules indicate what part of the message the content string
should be checked against.\Mar{Lucas, vc. tem dados para dar suporte a
  isto? ou seja, quantas options do tipo ``content'' possuem
  modificadores.} \Luc{conseguirei esses dados} For example, the
creator of the rule may want to indicate that a specific string be
present in the header of an HTTP request message. Content
\emph{modifiers} serve this purpose. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} states the requirement that the string ``keep-alive''
be present in the header of the message. (A keep-alive connection
attribute indicates that a single TCP connection should remain open
for multiple HTTP requests/responses.)  Figure~\Fix{adicione pequena
  figura...verifique se ainda faz sentido depois do que falei
  acima...}

@@

\tname\ uses the frequency of content modifiers to discriminate more
or less likely rules. For example, historical data indicates that
content options that use the \CodeIn{http\_connection} modifier are
much less likely than content options that use the modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the histogram of content modifiers for http. Although content
modifiers are applicable to rules of any protocol in Suricata,
currently, \tname\ only supports tcp and http. These are the protocols
involving ~80\% of the attacks.

@@

%\Mar{what about ``none'' (no modifier)?}}
%  \Luc{Uma regra pode sim ter contents sem modificadores, mas isso n
    %% vem ao caso aqui. Nessa heuristica toda string do payload vai ter
    %% associada a ela os modificadores q podem acessar a sua
    %% posição. Portanto, não vai existir o caso de não haver
    %% modificador. Atenção: isso não quer dizer q as regras q geramos
    %% incluem os modificadores, elas n incluem.

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''}. The contribution of this option to the score of
the corresponding rule would be $0.5$
(=$\frac{10,111}{10,111*1}$). The value of $M_i$ for this case is 1 as
the field of the HTTP message \CodeIn{Request URI}---holding the
string ``/HtmlAdaptor''---can be accessed through the modifier
\CodeIn{http\_uri}. Conceptually, the relatively high heuristic value
given to this option indicates that this is a reasonable option
because the string used in the content option (``/HtmlAdaptor'') is
stored in a specific field of the HTTP message (\CodeIn{Request URI})
and can be accessible through content modifiers
(\CodeIn{http\_uri}). For the option \CodeIn{content: ``keep-alive''},
either \CodeIn{http\_header} or \CodeIn{http\_coxnnection} could be
used as modifier to locate the string ``keep-alive'' stored in the
field \CodeIn{Connection} (see Figure~\Fix{xxx}). Consequently, for
this case, the contribution of modifiers to this content option would
be $0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
used in a rule, the aggregated contribution of this heuristic function
would be 0.35 (=(0.2+0.5)/2).

>>>>

\subsubsection{Frequency of content modifiers.} As previously
mentioned, \percRulesWithContent\ of Suricata rules involve content
options. Furthermore, we observed---by analyzing the dataset of
existing rules (see Section~\ref{sec:dataset-benign})---that often
these rules indicate what part of the message the content string
should be checked against.\Mar{Lucas, vc. tem dados para dar suporte a
  isto? ou seja, quantas options do tipo ``content'' possuem
  modificadores.} \Luc{conseguirei esses dados} For example, the
creator of the rule may want to indicate that a specific string be
present in the header of an HTTP request message. Content
\emph{modifiers} serve this purpose. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} states the requirement that the string ``keep-alive''
be present in the header of the message. (A keep-alive connection
attribute indicates that a single TCP connection should remain open
for multiple HTTP requests/responses.)
Figure \Fix{xxx}

@@

\tname\ uses the frequency of content modifiers to discriminate rules
more or less likely to reflect actual rules. For example, historical
data indicates that content options that use the
\CodeIn{http\_connection} modifier are much less likely than content
options that use the modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the histogram of content modifiers for http. Although content
modifiers are applicable to rules of any protocol supported by
Suricata, currently, \tname\ only supports tcp and http. These are the
protocols involving ~80\% of the rules from our dataset.

@@

%\Mar{what about ``none'' (no modifier)?}}
%  \Luc{Uma regra pode sim ter contents sem modificadores, mas isso n
    %% vem ao caso aqui. Nessa heuristica toda string do payload vai ter
    %% associada a ela os modificadores q podem acessar a sua
    %% posição. Portanto, não vai existir o caso de não haver
    %% modificador. Atenção: isso não quer dizer q as regras q geramos
%% incluem os modificadores, elas n incluem.
%\Luc{Inclui esse modificador pq ele tambem pode acessar o campo Request URI. Isso eh um problema? O http\_uri eh o modificador mais comum, mas o fato de haver o modificador http\_request\_line que tambem accessa a mesma posição faz com que o fitness maximo seja 0.5, pois a ocorrencia de http\_request\_line eh irrelevante no calculo.}

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''}. The contribution of this option to the score of
the rule that uses it would be $0.5$
(=$\frac{10,111+11}{10,111*2}$). The value of $M_i$ for this case is 2
as the field of the HTTP message \CodeIn{Request URI}---holding the
string ``/HtmlAdaptor''---can be accessed through the modifiers
\CodeIn{http\_uri} and \CodeIn{http\_request\_line}. The relatively
high heuristic value given to this option suggests it is a plausible
option to be used in rules---the string used in the content option
(``/HtmlAdaptor'') is stored in a specific field of the HTTP message
(\CodeIn{Request URI}) and can be accessed by two content modifiers,
namely (\CodeIn{http\_uri} and \CodeIn{http\_request\_line}). For the
option \CodeIn{content: ``keep-alive''}, either the modifier
\CodeIn{http\_header} or the modifier \CodeIn{http\_connection} could
be used to locate the string ``keep-alive'' stored in the field
\CodeIn{Connection} (see Figure~\Fix{xxx}). Consequently, for this
case, the contribution of modifiers to this content option would be
$0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
used in a rule, the aggregated contribution of this heuristic function
would be 0.35 (=(0.2+0.5)/2).
